---
title: "Create Dorling Cartogram"
output: html_document
---

```{r setup}

if(!require("cartogram")) install.packages("cartogram")
if(!require("tidycensus")) install.packages("tidycensus")


# Required packages
library(tidyverse)
library(tidycensus)
library(sf)
library(cartogram)

# Set options
options(tigris_use_cache = TRUE)
```

```{r get-data}
# Set your Census API key here
census_api_key("9330018ae4a2bd20da22ced1281145d9405dc349")

# Get county population data from 2020 Census
fips_pop <- get_decennial(
  geography = "county",
  variables = "P1_001N",  # Total population from 2020 Census
  year = 2023,
  cache_table = TRUE
) |>
  select(-variable) |>
  rename(
    fips = GEOID,
    name = NAME,
    pop = value
  ) |>
  mutate(
    fips = as.numeric(fips),  # Convert fips to numeric
    prop_pop = pop/sum(pop)
    )

# Get county geometries
fips_map <- get_decennial(
  geography = "county",
  variables = "P1_001N",
  year = 2023,
  geometry = TRUE,
  shift_geo = FALSE,
  cache_table = TRUE
) |>
  select(GEOID, NAME, geometry) |>
  rename(fips = GEOID, name = NAME) |>
  mutate(fips = as.numeric(fips))  # Convert fips to numeric
```

```{r create-dorling}
# Combine data and create initial sf object
counties_sf <- fips_map |>
  left_join(fips_pop, by = c("fips", "name")) |>
  # Transform to Albers equal area projection
  sf::st_transform(crs = 2163)

# Create the Dorling cartogram
county_dorling <- cartogram_dorling(
  x = counties_sf,
  weight = "prop_pop",
  k = 0.2,          # Adjust this to control spacing
  itermax = 100     # Increase for better convergence
)

# Filter out Alaska, Hawaii, and territories
county_dorling_filtered <- county_dorling |>
  filter(!str_detect(name, "Alaska|Hawaii|Puerto|Guam"))
```

```{r export}
# Export as GeoJSON
st_write(
  county_dorling_filtered,
  "county_dorling_ct.geojson",
  delete_dsn = TRUE  # Overwrites existing file
)

# Export as TopoJSON using geo2topo (requires mapshaper installed)
# This creates a more compact file
if(!require("rmapshaper")) install.packages("rmapshaper")
library(rmapshaper)
ms_write(
  county_dorling_filtered,
  "county_dorling_.topojson",
  force = TRUE
)
```

The key differences from the original code:
1. Only gets population data (not racial demographics)
2. Uses same cartogram settings but exports the result
3. Produces both GeoJSON and TopoJSON formats
4. Maintains all county attributes in the output files

To use this:
1. Save it as a .Rmd file
2. Make sure you have all required packages installed
3. Set your Census API key
4. Run all chunks

The resulting files will contain the Dorling cartogram geometry with:
- Properly spaced, non-overlapping bubbles
- Population-proportional sizing
- County FIPS codes and names preserved
- Continental US only

Would you like me to adjust any of the parameters (like the spacing factor `k` or number of iterations)?